Charity analysis

### 2022-07-20 Shiny/Test/Flexdashboard rows.Rmd
Letting styler do it's thing to make everything more readable.

```{r setup, include=FALSE}
# Loading packages using pacman to ensure they're all installed
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
       flexdashboard, MetBrewer, fontawesome, htmltools)
# Using fontawesome for more up-to-date fontawesome icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

@@

### Balance
```{r}

@@

#---- Data frames --------------------------------------------------------------      
    # creating data frame from html file
    htmldata <- reactive({
      read_proff8(input$html$datapath,
                  as.character(input$start), as.character(input$end))
      })

@@

    # Creating KPI_admin value
    KPI_admin <- reactive({
      exceldata() |> 
           mutate(
      KPI_admin =
        str_c( # Using str_c() to concatenate value and % sign
         format( # shortening(lengthening) the resulting value
           # calculating the value
          (expenses_administrative / expenses_program_services *100), digits = 3
          ), "%"
         )
      ) |> 
      filter(year == max(year)) |> # selecting latest available year
       select(KPI_admin) |> # dropping all other variables
        str_flatten() # turning into a single string
         })

    # Creating Fundraising ROI value
    KPI_fundraising_num <- reactive({
      exceldata() |> 
          mutate(
            # Calculating the value as a percentage
            KPI_fundraising_num =
              (revenue_private_contributions +
                 revenue_membership_fees - expenses_fundraising) /
              expenses_fundraising * 100) |> 
        # most recent year
          filter(year == max(year)) |> 
        # just want the single number
          select(KPI_fundraising_num) |> 
          pull()
    })

@@

    # Creating value box for fundraising ROI
      # Unsure about colour levels
    output$KPI_box2 <- renderValueBox({
      valueBox(
        # shortening(lengthening) the value with format()
        # Using str_c() to concatenate value and % sign;
        value = str_c(format(KPI_fundraising_num(), digits = 3), "%"),
        caption = "Fundraising ROI",
        # Colour depends on value
        color = if (KPI_fundraising_num() < 100) {"danger"
          } else if (KPI_fundraising_num() < 200) {"warning"
            } else {"success"},
        icon = "fas fa-donate"
      )
    })

    # Duplicate value boxes for testing
    output$KPI_box3 <- renderValueBox({
      valueBox(
        value = str_c(format(KPI_fundraising_num(), digits = 3), "%"),
        caption = "success",
        color = "success",
        icon = "fas fa-coffee"
      )
    })

    output$KPI_box4 <- renderValueBox({
      valueBox(
        value = KPI_admin(),
        caption = "info",
        color = "info",
        icon = "fas fa-skull-crossbones"
    )})

    output$KPI_box5 <- renderValueBox({
      valueBox(
        value = KPI_admin(),
        caption = "warning",
        color = "warning",
        icon = "fas fa-battery-empty"
    )})

#---- Plots --------------------------------------------------------------------    
#----CEO compensation-----------------------------------------------------------
    CEO_plot <- reactive(
      htmldata() |> 
        select(
          year, lederlonn, leder_annen_godtgjorelse
         ) |> 
        # Compensation in ordinary numbers
        mutate(
          lederlonn = lederlonn * 1000,
          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
        ) |> 
        # renaming for the purposes of the plot
        rename( 
          Year = year,
          Salary = lederlonn,
          `Other compensation` = leder_annen_godtgjorelse
        ) |> 
        # pivoting to join the two values to use fill
        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
        ggplot(aes(Year, value, fill = Compensation)) +
        geom_col() +
        scale_y_continuous(
          # Checking for uniform currency code
        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
                        # Using the currency code if uniform
                    label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
                        # alternative prefix if multiple currencies in the data
                    label_dollar(prefix = "? "))
         ) +
        # pretty colours
        scale_fill_met_d("Troy") + 
        labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
             # Checking again for currency; including warning if not uniform
        subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
                         "",
                         "Note: not all values are denominated in the same currency"))
        )

@@

#----Annual result---------------------------------------------
    arsresultat_plot <- reactive({
      htmldata() |>
      ggplot(aes(year, arsresultat, fill = annual_profit)) +
      geom_col() +
      scale_y_continuous(
        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
                        # Using the currency code if uniform
                    label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
                        # alternative prefix if multiple currencies in the data
                    label_dollar(prefix = "? "))
        ) +
        # slightly nicer colours + ensuring positive values are always green
      scale_fill_met_d("Tsimshian",
                   direction = ifelse(mean(htmldata()$annual_profit, na.rm = TRUE) == 1,-1, 1)) +
      theme(legend.position = "none") +
      labs(x = "Year", y = "", title = "Annual result (in thousands)",
             # Checking again for currency; including warning if not uniform
        subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
                         "",
                         "Note: not all values are denominated in the same currency"))
    })

    output$arsresultat_plotly <- renderPlotly(
      ggplotly(arsresultat_plot())
    )
#----Balance--------------------------------------------------------------------

})
```

```{r functions, include=FALSE}
# Function to read html file
read_proff8 <- function(file, start, end){
    read_html(file, encoding = "UTF-8") |> 
      html_node("#inner-frame") |> 
      html_table() |> 
      select( # removing empty column
        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
      mutate( # removing whitespace
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |> 
      filter( # removing duplicate table - no idea why it's necessary
        !row_number() > 176
      ) |>
      filter( # removing duplicate rows
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |> 
      pivot_longer( # tidying data
        all_of(end):all_of(start), names_to = "year"
        ) |> # using all_of()
      mutate( # changing to real NAs and turning years into numbers
        value = na_if(value, "-"),
        year = as.integer(year), 
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |> 
      fill( # filling the currency column
        valutakode, .direction = "updown"
      ) |> 
      filter( # removing dates and redundant currency
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode") |> 
      mutate( # removing whitespace in numbers
        value = str_replace_all(value, "\\s", ""),
        value = as.numeric(value),
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |> 
      distinct() |> # removing any remaining duplicate rows
      pivot_wider( # pivoting
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |> 
      clean_names() |> # tidying names
      select(-lederlonn_i_hele_1000,
             -resultatregnskap_i_hele_1000,
             -balanseregnskap_i_hele_1000) |> # removing headings
      rename( # making it clear what lonn refers to
        lederlonn = lonn
      ) |> 
      # Creating new variables for plots
      mutate(
        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
      ) |> 
      arrange(desc(year)) # arranging by year
}

# Function to read Excel file
widen_excel <- function(file){
    read_excel(file) |> 
      unite( # uniting Category and Name to deal with multiple 'Other'
        entry, c(Category, Name), sep = "_"
      ) |> 
      select(-`Sub-category`) |> # deleting Sub-category
    # pivoting
      pivot_wider(names_from = entry, values_from = Value) |> 
      clean_names() # cleaning names
}
```

>>>>

```{r setup, include=FALSE}
# Loading packages using pacman to ensure they're all installed
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
# Using fontawesome for more up-to-date fontawesome icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

@@

### Balance
```{r}
plotlyOutput("balance_plotly")
```

@@

  #---- Data frames --------------------------------------------------------------      
  # creating data frame from html file
  htmldata <- reactive({
    read_proff8(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

@@

  # Creating KPI_admin value
  KPI_admin <- reactive({
    exceldata() |>
      mutate(
        KPI_admin =
          str_c( # Using str_c() to concatenate value and % sign
            format( # shortening(lengthening) the resulting value
              # calculating the value
              (expenses_administrative / expenses_program_services * 100),
              digits = 3
            ), "%"
          )
      ) |>
      filter(year == max(year)) |> # selecting latest available year
      select(KPI_admin) |> # dropping all other variables
      str_flatten() # turning into a single string
  })

  # Creating Fundraising ROI value -- transmute?
  KPI_fundraising_num <- reactive({
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (revenue_private_contributions +
            revenue_membership_fees - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull()
  })

@@

  # Creating value box for fundraising ROI
  # Unsure about colour levels
  output$KPI_box2 <- renderValueBox({
    valueBox(
      # shortening(lengthening) the value with format()
      # Using str_c() to concatenate value and % sign;
      value = str_c(format(KPI_fundraising_num(), digits = 3), "%"),
      caption = "Fundraising ROI",

      # Colour depends on value
      color = if (KPI_fundraising_num() < 100) {
        "danger"
      } else if (KPI_fundraising_num() < 200) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-donate"
    )
  })

  # Duplicate value boxes for testing
  output$KPI_box3 <- renderValueBox({
    valueBox(
      value = str_c(format(KPI_fundraising_num(), digits = 3), "%"),
      caption = "primary",
      color = "primary",
      icon = "fas fa-coffee"
    )
  })

  output$KPI_box4 <- renderValueBox({
    valueBox(
      value = KPI_admin(),
      caption = "info",
      color = "info",
      icon = "fas fa-skull-crossbones"
    )
  })

  output$KPI_box5 <- renderValueBox({
    valueBox(
      value = KPI_admin(),
      caption = "warning",
      color = "warning",
      icon = "fas fa-battery-empty"
    )
  })

  #---- Plots --------------------------------------------------------------------    
  #----CEO compensation-----------------------------------------------------------
  CEO_plot <- reactive({
    htmldata() |>
      # preparing for pivoting
      select(
        year, lederlonn, leder_annen_godtgjorelse
      ) |>
      # Compensation in ordinary numbers
      mutate(
        lederlonn = lederlonn * 1000,
        leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
      ) |>
      # renaming for the purposes of the plot
      rename(
        Year = year,
        Salary = lederlonn,
        `Other compensation` = leder_annen_godtgjorelse
      ) |>
      # pivoting to join the two values to use fill
      pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
      ggplot(aes(Year, value, fill = Compensation)) +
      geom_col() +
      scale_y_continuous(
        # Checking for uniform currency code
        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
          # Using the currency code if uniform
          label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
          # alternative prefix if multiple currencies in the data
          label_dollar(prefix = "? ")
        )
      ) +
      # pretty colours
      scale_fill_met_d("Troy") +
      labs(
        x = "Year", y = "Total compensation", title = "Annual CEO compensation",
        # Checking again for currency; including warning if not uniform
        subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
          "",
          "Note: not all values are denominated in the same currency"
        )
      )
  })

@@

  #----Annual result---------------------------------------------
  arsresultat_plot <- reactive({
    htmldata() |>
      ggplot(aes(year, arsresultat, fill = annual_profit)) +
      geom_col() +
      scale_y_continuous(
        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
          # Using the currency code if uniform
          label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
          # alternative prefix if multiple currencies in the data
          label_dollar(prefix = "? ")
        )
      ) +
      # slightly nicer colours + ensuring positive values are always green
      scale_fill_met_d("Tsimshian",
        direction =
          ifelse(mean(htmldata()$annual_profit,
            na.rm = TRUE
          ) == 1, -1, 1)
      ) +
      theme(legend.position = "none") +
      labs(
        x = "Year", y = "", title = "Annual result (in thousands)",
        # Checking again for currency; including warning if not uniform
        subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
          "",
          "Note: not all values are denominated in the same currency"
        )
      )
  })

  output$arsresultat_plotly <- renderPlotly(
    ggplotly(arsresultat_plot())
  )
  #----Balance--------------------------------------------------------------------    
  balance_plot <- reactive({
    htmldata() |>
      ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
      geom_col() +
      scale_y_continuous(
        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
          # Using the currency code if uniform
          label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
          # alternative prefix if multiple currencies in the data
          label_dollar(prefix = "? ")
        )
      ) +
      # slightly nicer colours + ensuring positive values are always green
      scale_fill_met_d("Tsimshian",
        direction = ifelse(
          mean(htmldata()$balance_positive,
            na.rm = TRUE
          ) == 1, -1, 1
        )
      ) +
      theme(legend.position = "none") +
      labs(
        x = "Year", y = "", title = "Balance (in thousands)",
        # Checking again for currency; including warning if not uniform
        subtitle =
          ifelse(length(unique(htmldata()$valutakode)) == 1,
            "",
            "Note: not all values are denominated in the same currency"
          )
      )
  })

  output$balance_plotly <- renderPlotly(
    ggplotly(balance_plot())
  )
})
```

```{r functions, include=FALSE}
# Function to read html file
read_proff8 <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    select( # removing empty column
      REGNSKAPSPERIODE:all_of(start)
    ) |> # using all_of()
    mutate( # removing whitespace
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    filter( # removing duplicate table - no idea why it's necessary
      !row_number() > 176
    ) |>
    filter( # removing duplicate rows
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    pivot_longer( # tidying data
      all_of(end):all_of(start),
      names_to = "year"
    ) |> # using all_of()
    mutate( # changing to real NAs and turning years into numbers
      value = na_if(value, "-"),
      year = as.integer(year),
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill( # filling the currency column
      valutakode,
      .direction = "updown"
    ) |>
    filter( # removing dates and redundant currency
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate( # removing whitespace in numbers
      value = str_replace_all(value, "\\s", ""),
      value = as.numeric(value),
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    distinct() |> # removing any remaining duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    clean_names() |> # tidying names
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> # removing headings
    rename( # making it clear what lonn refers to
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    arrange(desc(year)) # arranging by year
}

# Function to read Excel file
widen_excel <- function(file) {
  read_excel(file) |>
    unite( # uniting Category and Name to deal with multiple 'Other'
      entry, c(Category, Name),
      sep = "_"
    ) |>
    select(-`Sub-category`) |> # deleting Sub-category
    # pivoting
    pivot_wider(names_from = entry, values_from = Value) |>
    clean_names() # cleaning names
}
```
